The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2015-03-26 draft.tex
made more corrections

Finally, \py\ originally modelled all bound-bound processes as transitions
within a simple two-level atom \cite[e.g.][]{mihalas}. 
This framework was used for the treatment of line transfer and also
for the line heating and cooling calculations. 
The approximation works reasonably well for resonance  
lines, such as \civfull, in which the lower level is the ground state.  
However, it is a poor approximation for many other
transitions, particularly those where the upper level
is primarily populated from above. Thus an improved method for
estimating excited level populations and simulating line transfer is
needed in order to model recombination lines and continua.

@@

In our implementation of the macro-atom approach, we also explicitly
take into account the photoionization and collisional ionization rates
between a lower level, $l$, and the continuum (or, in the case of ions
with more than one bound electron, the ground state of the upper ion),
$\kappa$,
\begin{equation}
{\cal R}_{l \kappa}= n_l \int_{\nu_0}^{\infty} \frac{ 4 \pi J_{\nu}
  \sigma_{\nu}}{h \nu} d\nu + C_{l \kappa} n_l n_e.
\end{equation}
Here, $\sigma_{\nu}$ is the photoionization cross section, and $J_{\nu}$
is the mean intensity. The corresponding recombination rate is given
by 
\begin{equation}
{\cal R}_{\kappa l} = \alpha_{\kappa l} n_{\kappa} n_e + C_{\kappa l}
n_\kappa n_e, \\
\end{equation}
where $\alpha_{\kappa l}$ is the radiative recombination coefficient
to level $l$. This treatment means that radiative and collisional
rates to and from all levels are considered when calculating both the
ionization state and the level populations, although we neglect 
ionization directly to excited levels of the upper ion. The
\cite{vanregemorter} approximation is used for collisional
transitions. This means that collisions between radiatively
forbidden transitions are not taken into account when one 
splits levels according to $l$ and $J$ quantum numbers, as well
as principal quantum number, $n$ (as we have done with He~\textsc{i}; 
see section~\ref{sec:data}).
Although this is, in general, a poor approximation, 
it is a second order effect in the physical regime where 
recombination lines are formed. 
This is because bound-free processes are 
dominant in determining level populations and emissivities.
We have verified that this is indeed the case in the He~\textsc{i}
emission regions in our model.

@@

Our main goal is to test whether the type of disk wind model that has
been successful in explaining the UV spectra of CVs could also have a
significant impact on the optical continuum and emission line spectra
of these systems. In order to set a benchmark, we therefore begin by
investigating one of the fiducial CV wind models that was used by SV93
and LK02 to simulate the UV spectrum of a typical high-state
system. The specific parameters for this model (model A) are listed in
Table~1. A key point is that the wind mass-loss rate in this model is
set to 10$\%$ of the accretion rate through the disk. 
The sensitivity to some of these parameters is discussed further in
section 5.

@@

@@

>>>>

Finally, \py\ originally modelled all bound-bound processes as transitions
within a simple two-level atom \cite[e.g.][]{mihalas}. 
This framework was used for the treatment of line transfer and also
for the line heating and cooling calculations (see LK02). 
The approximation works reasonably well for resonance  
lines, such as \civfull, in which the lower level is the ground state.  
However, it is a poor approximation for many other
transitions, particularly those where the upper level
is primarily populated from above. Thus an improved method for
estimating excited level populations and simulating line transfer is
needed in order to model recombination lines and continua.

@@

In our implementation of the macro-atom approach, we also explicitly
take into account the photoionization and collisional ionization rates
between a lower level, $l$, and the continuum (or, in the case of ions
with more than one bound electron, the ground state of the upper ion),
$\kappa$,
\begin{equation}
{\cal R}_{l \kappa}= n_l \int_{\nu_0}^{\infty} \frac{ 4 \pi J_{\nu}
  \sigma_{\nu}}{h \nu} d\nu + C_{l \kappa} n_l n_e.
\end{equation}
Here, $\sigma_{\nu}$ is the photoionization cross section, and $J_{\nu}$
is the mean intensity. The corresponding recombination rate is given
by 
\begin{equation}
{\cal R}_{\kappa l} = \alpha_{\kappa l} n_{\kappa} n_e + C_{\kappa l}
n_\kappa n_e, \\
\end{equation}
where $\alpha_{\kappa l}$ is the radiative recombination coefficient
to level $l$. This treatment means that radiative and collisional
rates to and from all levels are considered when calculating both the
ionization state and the level populations, although we neglect 
ionization directly to excited levels of the upper ion. The
\cite{vanregemorter} approximation is used for collisional
transitions. This means that collisions between radiatively
forbidden transitions are not taken into account when one 
splits levels according to $l$- and $s$-subshells, as well
as principal quantum number, $n$ (as we have done with He~\textsc{i}; 
see section~\ref{sec:data}). Although this is, in general, 
a poor approximation, it is a second order effect in the physical 
regime where recombination lines are formed. This is because bound-free processes are 
dominant in determining level populations and emissivities.
We have verified that this is indeed the case in the He~\textsc{i}
emission regions in our model.

@@

Our main goal is to test whether the type of disk wind model that has
been successful in explaining the UV spectra of CVs could also have a
significant impact on the optical continuum and emission line spectra
of these systems. In order to set a benchmark, we therefore begin by
investigating one of the fiducial CV wind models that was used by SV93
and LK02 to simulate the UV spectrum of a typical high-state
system. The specific parameters for this model (model A) are listed in
Table~1. A key point is that the wind mass-loss rate in this model is
set to 10$\%$ of the accretion rate through the disk. The inner edge of
the wind ($r_{min}$) is set to $4~R_{WD}$ following SV93. 
The sensitivity to some of these parameters is briefly discussed in
section~5.

@@

This model demonstrated that one can achieve a more
realistic optical spectrum by altering just two kinematic parameters. 
However, it may also be possible to achieve this with
other free parameters such as $\dot{M}_{wind}$, the opening angles of the wind and the 
inner and outer launch radii. For example, increasing the mass loss rate of the wind
would increase the amount of wind emission, as well as lowering the ionization 
parameter and increasing optical depth through the wind. 
Wider launch radii and opening angles would lead to a larger emitting volume, but
this would be moderated by the decrease in density for a fixed mass-loss rate. 
We also note that the inner radius of $4~R_{WD}$ adopted by SV93 
particularly affects the emergent spectrum seen at inclinations $<\theta_{min}$ as 
the inner disk is uncovered. A wind which extends right up to the WD would therefore
have more absorption in the resonance lines and a less pronounced Balmer absorption edge
at low inclinations. We suggest that future work should incorporate a treatment of clumping
and investigate the effect of these parameters in more detail. If a wind 
really does produce the line and continuum emission seen in optical spectra of high-state CVs, then
understanding the true mass loss rate and geometry of the outflow is clearly important.

@@

% \section{Sensitivity to Free Parameters}
% \label{sec:free_params}

% We have demonstrated that one can achieve a more
% realistic optical spectrum with small changes to the wind prescription; 
% specifically by altering just two kinematic parameters which increase the emission measure
% of the wind plasma. However, it may also be possible to achieve this by altering 
% other free parameters such as $\dot{M}_W$, the opening angles of the wind and the 
% inner and outer launch radii. Increasing the mass loss rate of the wind or volume of 
% the emitting area would

% We plan to explore the effect of these parameters
